---
title: "Untitled"
output:
  html_document:
    highlight: textmate
    theme: lumen
    toc: yes
    toc_float: yes
    toc_depth: 2
    css: custom.css
editor_options: 
  chunk_output_type: console
---

```{r}
# check to make sure user has the needed packages; if missing, install for them 
list.of.packages <- c("tidyverse", "remotes", "cowplot", "RColorBrewer")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)


remotes::install_github("OJWatson/hrp2malaRia")
remotes::install_github("nickbrazeau/EMSI2019hrp2lab") # hacky way to download this tutorial as a package
remotes::install_github("DavisVaughan/furrr") # we are going to use this for speed later
library(hrp2malaRia)
library(EMSI2019hrp2lab)
library(furrr)
library(tidyverse)
library(RColorBrewer)
library(cowplot)

set.seed(42) # please don't run this line! 
```


# Scenario 1: Different Diagnostics 
Imagine that there are three populations with a constant EIR of 1 (reminder, EIR is the entomogical innoculation rate, or the number of infectious mosquito bites per person per year). In addition, all populations start with approximately 10% frequency of hrp2 deletions and hrp2 deleted and wildtype parasites have equal fitness. However, the populations differ by:  
  
**Population 1:**

* Only uses rapid diagnostic tests (RDTs) to diagnose malaria


**Population 2:**

* Uses RDTs to diagnose 50% of cases and microscopy for the other 50%

**Population 3:**

* Only uses microscopy to diagnose malaria


### Output
As noted above, the `hrp2_Simulation` function returns many results that allow for many different questions to be investigated. For the purpose of our excerise, we will focus on the increase of hrp2 frequency (`df$S.N.Dels`) over time (`df$S.Times`) as well as incidence (`df$S.Incidence`) over time (`df$S.Times`)
  
Let's run the code (once) and plot the result: 


```{r}
# Simulations take ~1 minute to run on my MacBook Pro (3.1 GHz Intel Core i7)
# Population 1

# as a side note, let's track how long this takes
start_time <- Sys.time()

senario1.pop1 <- hrp2malaRia::hrp2_Simulation(N=2000, 
                                                  years=20, 
                                                  strains.0 = 10,
                                                  rdt.nonadherence = 0, 
                                                  EIR=1/365, 
                                                  microscopy.use = 0, 
                                                  rdt.det = 0, 
                                                  fitness = 1
                                                  )
# Population 2
senario1.pop2 <- hrp2malaRia::hrp2_Simulation(N=2000, 
                                                  years=20, 
                                                  strains.0 = 10,
                                                  rdt.nonadherence = 0, 
                                                  EIR=1/365, 
                                                  microscopy.use = 0.5, 
                                                  rdt.det = 0, 
                                                  fitness = 1
                                                  )


# Population 3
senario1.pop3 <-  hrp2malaRia::hrp2_Simulation(N=2000, 
                                                   years=20, 
                                                   strains.0 = 10,
                                                   rdt.nonadherence = 0, 
                                                   EIR=1/365, 
                                                   microscopy.use = 1, 
                                                   rdt.det = 0, 
                                                   fitness = 1
                                                   )


end_time <- Sys.time()

sequentialtime <- start_time - end_time

# Now, let's plot these - all veriables that begin with S. are series variables collected over time. 
# let's plot the first population
plot(senario1.pop1$S.Times,senario1.pop1$S.N.Dels, xlab = "time (days)", ylab = "hrp2 deletion Frequency", ylim=c(0,1), col="red", type="l", lwd=2)
# and add the second population
lines(senario1.pop2$S.Times, senario1.pop2$S.N.Dels,lwd=2, col="blue")
# and add the third population
lines(senario1.pop3$S.Times, senario1.pop3$S.N.Dels,lwd=2, col="green")
# and add our legend
legend(2000, .25, legend=c("Only RDT use", "50% RDT use", "0% RDT Use"),
       col=c("red", "blue", "green"), lty=1, cex=0.8,
       box.lty=2, box.lwd=2)

```





### Advanced Content
Given that this is a stochastic model, there will be run-to-run variability of each model iteration. This variability is important in quantifying our uncertainty of the stochastic process that we are modeling.   
To get multiple runs of the model, you can simply copy and paste the code above (and rename the output objects [e.g. `r1.2`], as to not overwrite your previous results). Or, we can make a **map** dataframe, where we specify the different parameters we want for each run and use this dataframe as the "captain" calling the "plays".     

In this **map dataframe**, our parameter names will be columns and our parameter options will be cells with each row being its own respective model. Using a **map** file is a cornerstone of functional programming and a high-yield concept for reproducible research. Plus, as Dr. Nunn noted earlier, stochastic models are typically slower to run than deterministic models.  As a result, we now have a need for speed, which is made easier with a **map dataframe** that we can "loop" through. 
  
Why do you think stochastic models are typically slower to run than deterministic models? _____________________
  
  
#### Very Advanced Content
We are going to use the `furrr` package to take advantage of the multiple proccessors on your computer (e.g. parallelize the simulations). This is simply for speed. If you are having trouble on your personal computer, you can change the `furrr::future_pmap` commands below to `purrr::pmap` and will get the same result (albiet at a slower pace). 
  
  
Let's try it out below: 

```{r}

hrp2map <- data.frame(run = c(1,1,1, 2,2,2, 3,3,3), # note you could use rep here (this is for clarity)
                      N=2000, 
                      years=20, 
                      strains.0 = 10,
                      rdt.nonadherence = 0, 
                      EIR=1/365, 
                      microscopy.use = c(0,0.5,1, 0,0.5,1, 0,0.5,1),
                      rdt.det = 0,
                      fitness = 1)
```

Let's look at what we just made: 
```{r}
dplyr::glimpse(hrp2map)
```
OK, let's run our models and store them in our **map dataframe**. 


```{r}
# note, we have to tell R that scenario and populations aren't parameters that hrp2_Simulation will accept
future::plan(multiprocess)

start_time <- Sys.time()

hrp2map$results <- furrr::future_pmap(hrp2map[,c("N", "years", "strains.0", "rdt.nonadherence", "EIR", "microscopy.use", "rdt.det", "fitness")], 
            hrp2malaRia::hrp2_Simulation)

end_time <- Sys.time()
paralleltime <- start_time - end_time


```

Just to prove a point, check out how much faster your code ran in parallel versus sequentially:
```{r}
sequentialtime
paralleltime
```

Now let's visualize our plots of the proportion of _hrp2_ deleted parasites over time.
```{r, fig.width = 11, fig.height = 20}

# note, I wrote this function for you, it returns a ggplot
plot.scenario1.run1 <- plothrp2models(pop1 = hrp2map$results[[1]],
                                      pop2 = hrp2map$results[[2]],
                                      pop3 = hrp2map$results[[3]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use")) # you need to manually add the labels

# you can manually add a title like this too 
plot.scenario1.run1 <- plot.scenario1.run1 + ggtitle("hrp2 Simulation Scenario 1, Run 1")


# now consider run 2
plot.scenario1.run2 <- plothrp2models(pop1 = hrp2map$results[[4]],
                                      pop2 = hrp2map$results[[5]],
                                      pop3 = hrp2map$results[[6]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use"))
plot.scenario1.run2 <- plot.scenario1.run2 + ggtitle("hrp2 Simulation Scenario 1, Run 2")

# now consider run 3
plot.scenario1.run3 <- plothrp2models(pop1 = hrp2map$results[[7]],
                                      pop2 = hrp2map$results[[8]],
                                      pop3 = hrp2map$results[[9]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use"))
                                      
plot.scenario1.run3 <- plot.scenario1.run3 + ggtitle("hrp2 Simulation Scenario 1, Run 3")

# Plot them together
cowplot::plot_grid(plot.scenario1.run1, 
                   plot.scenario1.run2,
                   plot.scenario1.run3,
                   ncol=1)




```



** Questions for Simulation 1 with regards to _hrp2_ frequency**

* How does the difference in diagnostic modality affect the proportion of _hrp2_ deletions? __________________
* How much variation did you appreciate between runs? __________________________________

* Was there anything unusual about your plots? Why was it unusual? _______________________________

Try running the models for a sample size of 100,000 (like the ELife paper). Note, this will take some time but the larger N should elicit a clearer/more consistent pattern.    



Now let's visualize our plots of the parasite Incidence over time. Incidence can be defined as the number of susceptible individuals that acquire an infection (at time $t$). 

```{r, fig.width = 11, fig.height = 20}

# note, I wrote this function for you, it returns a ggplot
iplot.scenario1.run1 <- plotincmodels(pop1 = hrp2map$results[[1]],
                                      pop2 = hrp2map$results[[2]],
                                      pop3 = hrp2map$results[[3]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use")) 


iplot.scenario1.run2 <- plotincmodels(pop1 = hrp2map$results[[4]],
                                      pop2 = hrp2map$results[[5]],
                                      pop3 = hrp2map$results[[6]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use"))


iplot.scenario1.run3 <- plotincmodels(pop1 = hrp2map$results[[7]],
                                      pop2 = hrp2map$results[[8]],
                                      pop3 = hrp2map$results[[9]],
                                      labels = c("Only RDT use", "50% RDT use", "0% RDT use"))
                                      


# Plot them together
cowplot::plot_grid(iplot.scenario1.run1, 
                   iplot.scenario1.run2,
                   iplot.scenario1.run3,
                   ncol=1)



```

** Questions for Simulation 1 with regards to Incidence **

* How did difference in diagnostic modality affect the overall incidence of _P. falciparum_ malaria? __________________
* How are incidence and proportion of _hrp2_ deletions related in these populations (in expectation)? __________________________________

* Why does that matter? _______________________________



# Scenario 2: Different Fitness 
Again, imagine that there are three populations with a constant EIR of 1 (reminder, EIR is the entomogical innoculation rate, or the number of infectious mosquito bites per person per year). In addition, all populations start with approximately 10% frequency of hrp2 deletions and all populations only use RDTs for diagnostics. However, the populations differ by:  

**Population 1:**

* Fitness is equal between hrp2-deleted and wildtype parasites

**Population 2:**

* Fitness cost of hrp2-deletion is approximately 0.5 compared to wildtype parasites
  
**Population 3:**

* Fitness benefit of hrp2-deletion is 1.5 compared to wildtype parasites

```{r, fig.width = 11, fig.height = 20}
# going to overwrite our old hrp2map and make a new one

hrp2map <- data.frame(
                      run = c(1,1,1,2,2,2),
                      N=2000, 
                      years=20, 
                      strains.0 = 10,
                      rdt.nonadherence = 0, 
                      EIR=1/365, 
                      microscopy.use = 0,
                      rdt.det = 0,
                      fitness = c(0.5, 1, 1.5, 0.5, 1, 1.5)
                      )


hrp2map$results <- furrr::future_pmap(hrp2map[,c("N", "years", "strains.0", "rdt.nonadherence", "EIR", "microscopy.use", "rdt.det", "fitness")], 
            hrp2malaRia::hrp2_Simulation)

plot.scenario2.run1 <- plothrp2models(pop1 = hrp2map$results[[1]],
                                      pop2 = hrp2map$results[[2]],
                                      pop3 = hrp2map$results[[3]],
                                      labels = c("hrp2 Fitness 0.5", "hrp2 Fitness 1", "hrp2 Fitness 1.5")) 
plot.scenario2.run1 <- plot.scenario2.run1 + ggtitle("hrp2 Simulation Scenario 2, Run 1")


plot.scenario2.run2 <- plothrp2models(pop1 = hrp2map$results[[4]],
                                      pop2 = hrp2map$results[[5]],
                                      pop3 = hrp2map$results[[6]],
                                      labels = c("hrp2 Fitness 0.5", "hrp2 Fitness 1", "hrp2 Fitness 1.5")) 
plot.scenario2.run2 <- plot.scenario2.run2 + ggtitle("hrp2 Simulation Scenario 2, Run 2")

cowplot::plot_grid(plot.scenario2.run1, plot.scenario2.run2,ncol=1)




```


** Questions for Simulation 2 with regards to _hrp2_ frequency **

* How do the differences fitness affect _hrp2_ frequency? __________________
* Were these effects greater than the effects of the diagnostic modality (above)? __________________________________
  * Why (or why not) do you think that may be the case? ___________________________
  * What are the differenes in a selective pressure vs. fitness? _________________________
* Can a parasite that is less fit still propagate through the population (call forward to Dr. Mitchell/Dr. Nunn's lecture) ? _______________________________   


**_If you have time, consider plotting the incidence for these simulations and compare them to scenario 1_**.





# Scenario 3: Different EIR 
Again, imagine that there are three populations where all populations start with approximately 10% frequency of hrp2 deletions, hrp2-deleted and wildtype strains have the same fitness, and all populations only use RDTs for diagnostics. However, the populations differ by:  

**Population 1:**

* EIR is 1 

**Population 2:**

* EIR is 10
  
**Population 3:**

* EIR is 100

```{r, fig.width = 11, fig.height = 20}
# going to overwrite our old hrp2map and make a new one

hrp2map <- data.frame(scenario = "scnr3",
                      population = c("pop1", "pop2", "pop3", "pop1", "pop2", "pop3"),  
                      run = c(1,1,1,2,2,2),
                      N=2000, 
                      years=20, 
                      strains.0 = 10,
                      rdt.nonadherence = 0, 
                      EIR=c(1/365, 10/365, 100/365, 1/365, 10/365, 100/365),
                      microscopy.use = 0,
                      rdt.det = 0,
                      fitness = 1
                      )


hrp2map$results <- furrr::future_pmap(hrp2map[,c("N", "years", "strains.0", "rdt.nonadherence", "EIR", "microscopy.use", "rdt.det", "fitness")], 
           hrp2malaRia::hrp2_Simulation)

plot.scenario3.run1 <- plothrp2models(pop1 = hrp2map$results[[1]],
                                      pop2 = hrp2map$results[[2]],
                                      pop3 = hrp2map$results[[3]],
                                      labels = c("EIR 1", "EIR 10", "EIR 100")) 
plot.scenario3.run1 <- plot.scenario3.run1 + ggtitle("hrp2 Simulation Scenario 3, Run 1")


plot.scenario3.run2 <- plothrp2models(pop1 = hrp2map$results[[4]],
                                      pop2 = hrp2map$results[[5]],
                                      pop3 = hrp2map$results[[6]],
                                      labels = c("EIR 1", "EIR 10", "EIR 100")) 
plot.scenario3.run2 <- plot.scenario3.run2 + ggtitle("hrp2 Simulation Scenario 3, Run 2")

cowplot::plot_grid(plot.scenario3.run1, plot.scenario3.run2,ncol=1)




```

** Questions for Simulation 3 with regards to _hrp2_ frequency **

* How do the differences EIR affect _hrp2_ frequency? __________________
  * Do any of the models seem odd (e.g. are any "overwhelmed")? _________________________
* Compare these effects with the effects above (above)? __________________________________
  * Why do you a high EIR results in so few _hrp2_-deleted parasites? ___________________________
  * Is there a parameter that we could change in our simulation that would allow for a high EIR and still produce a high proportion of _hrp2_-deleted parasites? _________________________



Now let's visualize our plots of the parasite Incidence over time. Incidence can be defined as the number of susceptible individuals that acquire an infection (at time $t$).   

Without looking at the models below, which population is going to have the highest incidence? ________________ 


```{r, fig.width = 11, fig.height = 20}

iplot.scenario3.run1 <- plotincmodels(pop1 = hrp2map$results[[1]],
                                      pop2 = hrp2map$results[[2]],
                                      pop3 = hrp2map$results[[3]],
                                      labels = c("EIR 1", "EIR 10", "EIR 100")) 

iplot.scenario3.run2 <- plotincmodels(pop1 = hrp2map$results[[4]],
                                      pop2 = hrp2map$results[[5]],
                                      pop3 = hrp2map$results[[6]],
                                      labels = c("EIR 1", "EIR 10", "EIR 100")) 

cowplot::plot_grid(iplot.scenario3.run1, iplot.scenario3.run2, ncol=1)



```


